Further Evidence of Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2022-02-02 src/ms.tex
adding shifted models figure

@@

%Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff. For higher mass stars with MS lifetimes of 4 Gyr, it means that stars on the ridge should have a ~2 Gyr age spread. For stars with 10 Gyr MS lifetimes, it should be something close to a 5 Gyr age spread.

@@

@@

The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge. We note that this is a speculative scenario and the underlying physics are not known.

>>>>

\begin{figure*}
    \centering 
    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks-shifted.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks-shifted.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost-shifted.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost-shifted.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models (2-d histograms) presented in \citet{vanSaders2019}. Shown here are stars with \logg~$>$~4.1. Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation. The black symbol in each panel indicates the position of the Sun.}
    \label{fig:shifted}
\end{figure*}

@@

%This is a bit misleading. Stars reach Ro = 2 at very roughly halfway through the MS, and stay there until the MS turnoff. For higher mass stars with MS lifetimes of 4 Gyr, it means that stars on the ridge should have a ~2 Gyr age spread. For stars with 10 Gyr MS lifetimes, it should be something close to a 5 Gyr age spread.

@@

%JVS: I am not sure this is actually what you expect to see. If you go back and look at the Lanzafame and Spada paper from 2015, the spin down never stalls out entirely... there is always evolution in the period, even though it is slower than you would have predicted with a solid body model. In the solar mass stars, you never see something as stark as the Praesepe/NGC6811 overlap. You're also working in a regime where initial conditions, saturated spin down, and cluster age uncertainties can really get you. I'm not sure you can attribute the overlap in Fritzewski to core-envelope decoupling... it could also just be you are looking at clusters that are coeval within the age uncertainties (remember, you're dealing with isochrone ages that come from stars with convective cores at these ages, and they have *nasty* systematic shifts depending on what you assumed about the degree of convective core overshooting).  I would tread carefully here.

@@

\textbf{The data presented here provide evidence for the gap being caused by or related to the same phenomena as the short-period pile-up, namely, core-envelope coupling. The decline in photometric variability that coincides with the gap is likely not the cause of it, but may result from a period of altered or decreased magnetic activity that is connected with core-envelope coupling. Connections between the short-period pile-up, the decline in photometric variability, and the existence of the intermediate period gap may also be consistent with the suggestion discussed in \citet{Gordon2021} that the gap could be due to a period of accelerated spin-down immediately proceeding the stalling due to core-envelope coupling. In this picture, surface spin-down slows as angular momentum is transferred from the core to the envelope, which is accompanied by a decline in photometric variability. After some time, surface spin-down resumes, but at an accelerated rate such that stars cross the gap quickly and are rarely observed there. A temporary increase in photometric variability accompanies this period of post-stalling spin-down, before declining again as stars spin down to the long-period edge. We note that this is a speculative scenario and the underlying physics are not known.}

%JVS: I am confused by this picture. Why would the slow in apparent spin down make the photometric variability drop? The variability doesn't care about the spin-down rate, it cares about the instantaneous rotation (and maybe differential rotation). In this picture you are still stuck without a reason for why a more slowly rotating star should be more active than a more rapidly rotating one. The information you add here doesn't allow you to support or reject the speculation in Gordon 2021, so I would vote to make this speculation more modest, saying that the gap is perhaps an epoch of rapid spin down and leave it at that.
